---
title: "SetSolutionConcepts"
author: "Jochen Staudacher"
date: "6 March 2019"
output: pdf_document
---

# Set Solution Concepts and Allocation Properties

The package `CoopGame` provides the following five set-based
solution concepts:

* the imputation set 
* the core
* the core cover
* the reasonable set
* the Weber Set

For all five set-based solution concepts we provide

* a routine for computing the vertices of the corresponding set
for an arbitrary number $n \geq 2$ of players
* a routine for checking whether an allocation $x$ is
part of the corresponding set
* a routine for drawing the corresponding set for $3$ or
$4$ players (see chapter 6 of this vignette)

Note that both for checking whether an allocation $x$
belongs to a set solution concept or not and for drawing 
the set solution (in the case of $3$ or $4$ players)
the corresponding vertices are internally
always computed first.

## Available set solution concepts

### The imputation set
Given a TU game $v$ with $n$ players we call a vector 
$x \in \mathbb{R}^{n}$ **efficient** if 
$$ \sum_{i=1}^{n} x_{i} = v(N).$$
Frequently the set of efficient vectors is also called 
the set of preimputations, see e.g. the 
book by Peleg and Sudhölter [@PelegSudhoelter], p. 20.
In case $x \in \mathbb{R}^{n}$ is also **individually rational**
(from the point of view of every player $i$), i.e.
$$ v(\{i\}) \leq x_{i}$$
for all $i=1, \dots, n$, then we call $x$ an imputation.
Formally, we can specify the so-called imputation set $I(v)$ as
$$I(v)= \{ x \in \mathbb{R}^{n} \vert \sum_{i=1}^{n} x_{i} = v(N)
\land v(\{i\}) \leq x_{i} \}$$
According to our game property definitions from section 3 the
imputation set is empty unless $v$ is either essential or
degenerate. In the latter case $I(v)$ consists of a single point.
For further details on imputations see e.g. the book by 
Peleg and Sudhölter 
[@PelegSudhoelter], p.20, 
the book by Maschler, Solan and Zamir 
[@MaschlerSolanZamir], p. 674--677,
or the book by 
Osborne and Rubinstein [@OsborneRubinsteinBook], p. 278.  

In order to computate the vertices of the imputation set
we provide the function `imputationsetVertices`. The rows
of the matrix returned are the vertices of the imputation set.
In case the imputation set is empty an empty matrix is returned.
In order
to check whether an allocation $x$ is an imputation there
is the function `belongsToImputationset`.
```{r, collapse=TRUE}
library(CoopGame)
v <-c(0,0,0,1,1,0,1)
imputationsetVertices(v)
belongsToImputationset(c(0.7,0.3,0),v)
```

### The core 
Apart from the set of imputations the most prominent set
solution concept is the **core**. The concept was established in
the Ph.D. thesis by Gillies [@GilliesThesis] from 1953 and a 
seminal article by Aumann [@aumann1961core] from 1961. 
The idea of the core is to consider only those imputations $x$
which are also **coalitionally rational**, i.e.
$$ v(S) \leq \sum_{i \in S} x_{i}$$
for all nonempty coalitions $S \subseteq N$. More 
information on the core can e.g. be found in  
the book by Maschler, Solan and Zamir 
[@MaschlerSolanZamir], pp. 686--747,
the book by Osborne and Rubinstein 
[@OsborneRubinsteinBook], pp. 257--275, or 
the book by Peleg and Sudhölter 
[@PelegSudhoelter], pp. 27--49.

The function `coreVertices` lists the vertices of the core
as rows of a matrix.
In case the core is empty an empty matrix is returned.
The function `belongsToCore` checks whether an allocation $x$
is contained in the core or not.
```{r, collapse=TRUE}
library(CoopGame)
v <-c(0,0,0,1,1,0,3)
coreVertices(v)
belongsToCore(c(1.7,1.3,0),v)
```

### The core cover

The core cover was originally suggested 
by Tijs and Lipperts in [@tijs1982CoreCover]
as a core catcher. 
Given a TU game $v$ with $n$ players let $m$ denote
the minimal rights vector and $M$ the utopia payoff vector.
The core cover of the game $v$ consists of all
imputations $x$ satisfying
$$m(i) \leq x_{i} \leq M(i)$$
for all $i=1, \dots, n$. One can show that the core is always
a subset of the core cover, see e.g. the book by 
Branzei, Dimitrov and Tijs [@BranzeiBook], p. 21, 
or the book by Chakravarty, Mitra and Sarkar 
[@ChakravartyMitraSarkar], pp. 45--46, for more details.
Also note that the core cover is nonempty iff the game $v$
is quasi-balanced (see chapter 3 of this vignette).

We provide functions `coreCoreVertices` listing the 
vertices of the core cover
as rows of a matrix and 
`belongsToCoreCover` for checking whether an allocation $x$
is contained in the core cover or not.

### The reasonable set

The reasonable set was originally 
suggested by Milnor in [@milnor1953reasonable]
as another core catcher. 
The reasonable set of a TU game $v$ consists of all
imputations $x$ satisfying
$$v(\{i \}) \leq x_{i} \leq 
\max_{i: i \in S} (v(S) - v(S \backslash \{i \}) $$
for all $i=1, \dots, n$. One can show that the core
cover is always
a subset of the reasonable set.  
We refer to the book by 
Branzei, Dimitrov and Tijs [@BranzeiBook], p. 21, 
the book by Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], pp. 43--46, and 
the article by 
Gerard-Varet and Zamir [@gerard1987remarks] 
for more details 
concerning the reasonable set.

We provide functions `reasonableSetVertices` listing the 
vertices of the reasonable set
as rows of a matrix and 
`belongsToReasonableSet` for checking whether 
an allocation $x$
is contained in the reasonable set or not.

### The Weber Set

The Weber Set was originally 
suggested by Weber in [@Weber1988]
as another core catcher. 
The Weber Set is the convex hull of
the marginal vectors. 
Thus one can easily see that the Weber Set
is contained in the imputation set for
superadditive games.
We provide the Weber Set only for game vectors $v$
which are both nonnegative and monotonic.
One can show that the core is always
a subset of the Weber Set. We refer to the 
book by Peters [@PetersBook], chapter 18, 
for further details 
regarding the Weber Set.

We provide functions `weberSetVertices` listing the 
vertices of the Weber Set
as rows of a matrix and 
`belongsToWeberSet` for checking whether 
an allocation $x$
is contained in the Weber Set or not.

## Using `rcdd` for computations

All five set-based solution concepts yield convex polyhedra 
as results. 
We wish to stress that we never coded any vertex computations
ourselves, but relied on the available R packages
`rcdd` [@rcdd] and `geometry` [@geometry]. 
We wish to give particular praise to the library `cdd` for 
polyhedral computations 
by Komei Fukuda [@cddFukuda]. In turn, 
the availability of a package like 
`rcdd` [@rcdd] 
as an R-interface of the powerful library `cdd` underlines 
the power of the `R` ecosystem.

The imputation set, the core, the core cover
and the reasonable set are 
by their definitions implicitly specified
as so-called H-representations, i.e. 
intersections of halfspaces, of convex polyhedra. 
Computing the so-called V-representations, i.e. 
representing the polyhedron as the convex hull
of its vertices, is done via the package `rcdd`.
The Weber set is already defined as the convex
hull of the marginal vectors. Hence we only
need to use `rcdd` to remove any redundant
points from a list of potential 
Weber set vertices. This feature of removing redundant
points is also used to determine whether a
given allocation $x$ belongs to one of the 
five set solution concepts in question or not.
For further information on convex polyhedra we simply
refer to the books by Rockafellar [@rockafellar2015convex]
and Ziegler [@ziegler2012lectures].