---
title: "PointSolutionConcepts"
author: "Jochen Staudacher"
date: "6 March 2019"
output: pdf_document
---

# Point Solution Concepts

The package `CoopGame` prides itself with implementations 
of a very large amount of point valued solution concepts. 
Calling a specific point valued solution concept always 
works the same way, i.e. `nameOfPointValuedSolutionConcept(v)` 
where $v$ stands for the game vector of a cooperative game.

This chapter is divided into two parts: In the first part 
we will give an extremely brief and 
concise overview of the point 
valued solution concepts available. In the second part we 
will guide the readers through three examples of 
point solutions of TU games and another two examples 
of power index computations.


## Overview of point solution concepts in `CoopGame`

We distinguish the point solution concepts we implemented 
by solution concepts for general cooperative games and 
power indices for simple games. 
We tried to document our implementations of these 
solution concepts in detail and give at least one 
reference for each individual function. Every time there is 
more than one reference we checked carefully that 
definitions were not contradictory.

### Available point solution concepts for general cooperative games 

The following 19 point solution concepts for 
general cooperative games are available in `CoopGame`:

+ the **centroid of the core** provided by the function `centroidCore`
+ the **centroid of the core cover** provided by the function `centroidCoreCover`
+ the **centroid of the imputation set** provided by the function `centroidImputationSet`
+ the **centroid of the reasonable set** provided by the function `centroidReasonableSet`
+ the **centroid of the Weber set** provided by the function `centroidWeberSet`
+ the **disruption nucleolus** (see 
the article by Littlechild and Vaidya [@LittlechildVaidya]) 
provided by the function `disruptionNucleolus`
+ the **Gately point** (see 
the original paper by Gately [@GatelyPaper], 
our recent article [@StaudacherAnwander] for 
implementation details, the article by Littlechild and Vaidya 
[@LittlechildVaidya] or the book 
by Narahari [@NarahariBook], pp. 455--456) 
provided by the function `gatelyValue`
+ the **modiclus** (aka modified nucleolus, 
see the one of the two original papers by 
Sudhölter [@Sudhoelter1996], [@Sudhoelter1997] 
or the book by Peleg and Sudhölter 
[@PelegSudhoelter], pp. 147--182) 
provided by the function `modiclus`
+ the **normalized Banzhaf value** (see e.g. one of 
the articles by Gambarelli [@gambarelli2011banzhaf] or 
by Stach [@stach2017sub])
provided by the function `normalizedBanzhafValue`
+ the **nucleolus** (see the papers by 
Schmeidler [@schmeidler1969nucleolus], 
by Kohlberg [@kohlberg1971nucleolus], 
by Kopelowitz [@kopelowitz1967computation], 
by Megiddo [@megiddo1974nonmonotonicity]
or the book by Peleg and Sudhölter 
[@PelegSudhoelter], pp. 82--86) 
provided by the function `nucleolus`
+ the **per capita nucleolus** (see the 
original paper by Young 
[@Young1985]) provided by the function `perCapitaNucleolus`
+ the **prenucleolus** (see chapter 6 of 
the book by Peleg and Sudhölter [@PelegSudhoelter]) 
provided by the function `prenucleolus`
+ the **proportional nucleolus** (see the 
original paper by Young, Okada and Hashimoto 
[@YoungOkadaHashimoto]) 
provided by the function `proportionalNucleolus`
+ the **(normalized) Public Good value** (see the 
original paper by Holler and Li 
[@holler1995public]) provided 
by the function `publicGoodValue` 
+ the **(normalized) Public Help value Chi** (see 
the original paper by Bertini and Stach 
[@BertiniStach2015]) provided 
by the function `publicHelpChiValue` 
+ the **(normalized) Public Help value Theta** (see 
the original paper by Bertini and Stach 
[@BertiniStach2015]) provided 
by the function `publicHelpValue` 
+ the **Shapley value** (see 
the original paper by Shapley [@shapley1953value], 
the article by Bertini [@bertini2011shapley] or chapter 18 
of the book 
by Maschler, Solan and Zamir [@MaschlerSolanZamir]) provided 
by the function `shapleyValue`
+ the **simplified modiclus** (see 
the original article by Tarashnina [@Tarashnina2011]) 
provided by the function `simplifiedModiclus`
+ the **tau-value** (aka Tijs value, see 
the original paper by Tijs [@Tijs1981], 
the article by Stach 
[@stach2011tijs] or the book by 
Branzei, Dimitrov and Tijs [@BranzeiBook], p. 31)
provided by the function `tauValue`

Note that the above 19 solution concepts are all efficient, 
i.e. the sum of the components of the solution equals the 
value of the grand coalition $v(N)$.

In addition, `CoopGame` also provides five solution concepts 
for general cooperative games which are not efficient:

+ the **absolute Public Good value** (see 
the original paper by Holler and Li 
[@holler1995public]) provided 
by the function `absolutePublicGoodValue` 
+ the **absolute Public Help value Chi** (see 
the original paper by Bertini and Stach 
[@BertiniStach2015]) provided 
by the function `absolutePublicHelpChiValue` 
+ the **absolute Public Help value Theta** (see 
the original paper by Bertini and Stach 
[@BertiniStach2015]) provided 
by the function `absolutePublicHelpValue` 
+ the **Banzhaf value** (see 
the book by Peters [@PetersBook], pp. 367--370)
provided by the function `banzhafValue`
+ the **raw Banzhaf value** (see 
the book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], pp. 118--119)
provided by the function `rawBanzhafValue`

### Available power indices for simple games

The following seven power indices for 
simple games are available in `CoopGame`:

+ the **Deegan-Packel index** (see 
the original article by Deegan and Packel [@DeeganPackel] 
or the German book by Holler and Illing 
[@HollerIlling], pp. 323--324)
provided by the function `deeganPackelIndex`
+ the **Johnston index** (see 
the original paper by Johnston 
[@johnston1978measurement] 
or the book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], p. 124)
provided by the function `johnstonIndex`
+ the **normalized Banzhaf value** (see 
the article by Stach [@stach2011banzhaf] 
or the book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], pp. 118--119) 
provided by the function `normalizedBanzhafIndex`
+ the **Public Good index** (see the original 
papers by Holler [@Holler1982], by Holler and Packel  
[@HollerPackel] or the later article by Holler 
[@holler2011public]) provided 
by the function `publicGoodIndex` 
+ the **Public Help index Chi** (see the original 
papers by Bertini and Stach 
[@BertiniStach2015] and by Stach [@Stach2016]) provided 
by the function `publicHelpChiIndex` 
+ the **Public Help index Theta** (see 
the original paper by 
Bertini, Gambarelli and Stach [@BertiniGambarelliStach] or 
one of the later articles by Bertini and Stach 
[@BertiniStach2015] or Stach [@Stach2016]) provided 
by the function `publicHelpIndex` 
+ the **Shapley-Shubik index** (see 
the original paper 
by Shapley and Shubik [@shapley1954method], 
the article by Stach 
[@stach2011shapley] from 2011 
or the book by Peters [@PetersBook], 
pp. 156--159) provided 
by the function `shapleyShubikIndex`

Note that the above seven power indices are all efficient, 
i.e. the sum of the components of the solution equals $1$.

In addition, `CoopGame` also provides nine power indices 
for simple games which are not efficient:

+ the **Barua Chakravarty Sarkar index** (see 
the original paper by 
Barua, Chakravarty and Sarkar [@barua2012measuring] or the 
book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], pp. 120--123)
provided 
by the function `baruaChakravartySarkarIndex`
+ the **Coleman Power index of a Collectivity to Act** 
(see the original article by 
Coleman [@coleman1971control], the paper 
by Stach [@stach2011coleman] or 
the survey article 
by de Keijzer [@deKeijzer2008survey], p. 18) 
provided 
by the function `colemanCollectivityPowerIndex`
+ the **Coleman Initiative Power index** 
(see the original article by Coleman [@coleman1971control], 
the paper by Stach [@stach2011coleman], 
the survey article 
by de Keijzer [@deKeijzer2008survey], p. 18, or 
the book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], pp. 120--123) 
provided 
by the function `colemanInitiativePowerIndex`
+ the **Coleman Preventive Power index** 
(see the original article by Coleman [@coleman1971control], 
the paper by Stach [@stach2011coleman], 
the survey article 
by de Keijzer [@deKeijzer2008survey], p. 18, or
the book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], pp. 120--123) 
provided 
by the function `colemanPreventivePowerIndex`
+ the **König-Bräuninger index** (see one of the 
original articles by König and Bräuninger 
[@konig1998inclusiveness] and by 
Nevison, Zicht and Schöpke [@nevison1978naive] or 
the later paper by Bertini and Stach 
[@BertiniStach2015]) provided 
by the function `koenigBraeuningerIndex`
+ the **Nevison index** (see 
the original paper by Nevison [@nevison1979structural]) 
provided 
by the function `nevisonIndex`
+ the **non-normalized Banzhaf index** (see 
the paper by Stach [@stach2011banzhaf] from 2011 
or the book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], pp. 118--119) 
provided 
by the function `nonNormalizedBanzhafIndex`
+ the **Rae index** (see the original article 
by Rae [@rae1969decision] or 
the book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], p. 119--120) 
provided by the function `raeIndex`
+ the **raw Banzhaf index** (see 
the book by 
Chakravarty, Mitra 
and Sarkar [@ChakravartyMitraSarkar], p. 118--119) 
provided by the function `rawBanzhafIndex`

In the context of power indices this vignette is not 
the place to discuss any details concerning the 
concepts of I-Power and P-Power. Hence we simply wish 
to refer to the original paper by Felsenthal, 
Machover and Zwicker [@felsenthal1998bicameral] and 
the book by Felsenthal and 
Machover [@felsenthal1998measurement] for details. The 
two small overview articles [@felsenthal2011IPower] and 
[@felsenthal2011PPower] 
by Felsenthal and Machover from 2011 may serve as 
a very first introduction to these concepts.

## A few remarks on the implementation of the nucleolus and its derivatives 

We wish to stress that in our implementations of 
the nucleolus and its various derivatives, i.e. the 
prenucleolus, the per capita nucleolus, the proportional 
nucleolus, the disruption nucleolus, the modiclus and 
the simplified modiclus, we followed the paper 
by Guajardo and 
Jörnsten [@CommonMistakesNucleolus] very closely as to avoid 
the common mistakes in computing the nucleolus described 
in this article. In particular, we made sure to 
incorporate the
information on the dual values in the solution process 
as described in the paper 
by Guajardo and 
Jörnsten [@CommonMistakesNucleolus]. 
Our algorithmic approach for finding the nucleolus and its 
derivatives is nothing particularly fancy. We simply 
solve a series of linear programs following the 
technical report by Kopelowitz [@kopelowitz1967computation] from 
1967. Our solver for the individual linear programs is 
the revised dual simplex solver provided by the function `lpcdd` 
from the package `rcdd` [@rcdd]. We finally note that an 
earlier yet unpublished version of `CoopGame` employed the 
package `glpkAPI` [@glpkAPI], a low-level interface to 
the GLPK library [@makhorin2008glpk] by Makhorin, for 
the solution of linear programs. We refer to 
the M.Sc. thesis of the second author [@AnwanderMaster] 
from 2017 for 
details of this unpublished approach. 
Later studies of the first author showed that the 
revised dual simplex solver provided by the function `lpcdd` 
from the package `rcdd` [@rcdd] was much better suited to 
compute the nucleolus and its derivatives and lead to 
reimplementation before the first submission 
of `CoopGame` to CRAN.

## Using the point solution concepts in `CoopGame` by example 

We would now like to guide our readers through three 
classic examples of point solutions for cooperative games.

### The Shapley value for the inheritance problem due to Ibn Ezra (1146)

Our first example is taken from a paper by Robert Aumann 
from 2010 on 
"Some non-superadditive games, and their 
Shapley values, in the Talmud" [@aumann2010some].
As the first in a series of problems, 
Aumann investigates the inheritance problem due to Ibn Ezra
(1146) (see [@ibn18951146]), i.e. an 
inconsistent will, in that paper. 

A man with four sons dies. 
He leaves an estate worth 120 units of money. He
bequeathes 120 units of money to 
his first son, 60 units of money to his 
second son, 40 units of money to his third sum 
and 30 units of money to his youngest son. 
In order to analyze the problem of how to divide 
the estate, let us first introduce the 
estate $E=120$ and the four claims $c_{1}=120$, 
$c_{2}=60$, $c_{3}=40$ and $c_{4}=30$. We can now 
define a TU game $v$ as follows:

$$
v = \left\{
\begin{array}{rl}
\min(E, \max_{i \in S}c_{i}) & \text{if } \vert S \vert < 4,\\
E & \text{else,}
\end{array} \right. 
$$

```{r, collapse=TRUE}
library(CoopGame)
Aumann2010Example<-c(120,60,40,30,120,120,120,60,60,40,120,120,120,60,120)
shapleyValue(Aumann2010Example)
```

We confirm that the Shapley value of $v$ coincides with 
Ibn Ezra’s solution: The first son get $80 \frac{5}{6}$, 
the second son $20 \frac{5}{6}$,
the third son $10 \frac{5}{6}$ and 
the youngest son $7 \frac{1}{2}$ units of money. 
For more details, we refer to the article by 
Aumann [@aumann2010some] and the 1982 paper 
by O'Neill [@ONeill1982problem]. 

### The nucleolus for bankruptcy problems from the Babylonian Talmud
Our second set of examples revisits bankruptcy 
problems from the Babylonian Talmud from section 2.1, 
see the papers by Aumann and Maschler [@AumannMaschler1985] and 
Aumann [@Aumann2002]. We solve the problem of the three 
widows from the Babylonian Talmud for $E=100, 200$, and $300$ 
using our function `nucleolus`:

```{r, collapse=TRUE}
library(CoopGame)
v100 = bankruptcyGameVector(n=3,d=c(100,200,300),E=100)
nucleolus(v100)
v200 = bankruptcyGameVector(n=3,d=c(100,200,300),E=200)
nucleolus(v200)
v300 = bankruptcyGameVector(n=3,d=c(100,200,300),E=300)
nucleolus(v300)
```
We confirm that the nucleolus coincides with the solutions 
from the Babylonian Talmud in all three cases. For more 
details, e.g. on the principle of equal division of the 
contested sum we refer to the article by 
Aumann and Maschler [@AumannMaschler1985] and 
chapter 4 of the book by Gura and Maschler [@GuraMaschlerBook].

### Solving the TVA problem
We investigate the classic TVA problem 
introduced in section 2 of this vignette (see the 
original article by Ransmeier 
[@ransmeier1942tennessee]
on the Tennessee Valley Authority 
from 1942 for the history of the problem and the article 
by Straffin and Heaney [@StraffinHeaney] 
and the book by Driessen [@DriessenBook], p. 98, for 
computational results).

```{r, collapse=TRUE}
library(CoopGame)
TVACosts=c(163520,140826,250096,301607,378821,367370,412584)
(v <- costSharingGameVector(n=3, TVACosts))
TVACosts[1:3] - gatelyValue(v)
TVACosts[1:3] - shapleyValue(v)
TVACosts[1:3] - nucleolus(v)
TVACosts[1:3] - tauValue(v)
```
              
We confirm the results from the literature.
Note that for the TVA problem the results for the 
tau-value and the Gately point need to coincide as 
the game is semiconvex, see [@driessen1985tau].

## Using the power indices in `CoopGame` by example 

We would now like to guide our readers through two 
examples of applications of power indices.

### Examples on the failure of the donation property

Bertini, Freixas, Gambarelli and Stach point out in 
their article [@bertini2013comparing] that the 
normalized Banzhaf index, the Deegan-Packel index,
the Public Good index, the Johnston 
index and the Public Help index Theta lack the 
so-called donation property. We revisit the examples from 
table 1 on page 9 of that paper and along the way 
point out that the donation property also does not hold 
for the Public Help index Chi. We start with the 
5-player game from the article by 
Bertini, Freixas, Gambarelli and Stach [@bertini2013comparing]:

```{r, collapse=TRUE, eval=FALSE}
library(CoopGame)
v<-weightedVotingGameVector(n=5, w=c(6,4,1,1,1), q=9)
normalizedBanzhafIndex(v)
## [1] 0.47368421 0.36842105 0.05263158 0.05263158 0.05263158
deeganPackelIndex(v)
## [1] 0.375 0.250 0.125 0.125 0.125
publicGoodIndex(v)
## [1] 0.3333333 0.1666667 0.1666667 0.1666667 0.1666667
johnstonIndex(v)
## [1] 0.52777778 0.38888889 0.02777778 0.02777778 0.02777778
publicHelpIndex(v)
## [1] 0.28125 0.25000 0.15625 0.15625 0.15625
publicHelpChiIndex(v)
## [1] 0.3234568 0.3003086 0.1254115 0.1254115 0.1254115
#
# Now player 1 donates one vote to player 2
v<-weightedVotingGameVector(n=5, w=c(5,5,1,1,1), q=9)
normalizedBanzhafIndex(v)
## [1] 0.5 0.5 0.0 0.0 0.0
deeganPackelIndex(v)
## [1] 0.5 0.5 0.0 0.0 0.0
publicGoodIndex(v)
## [1] 0.5 0.5 0.0 0.0 0.0
johnstonIndex(v)
## [1] 0.5 0.5 0.0 0.0 0.0
publicHelpIndex(v)
## [1] 0.2857143 0.2857143 0.1428571 0.1428571 0.1428571
publicHelpChiIndex(v)
## [1] 0.3309524 0.3309524 0.1126984 0.1126984 0.1126984
```

Now for the example with 10 players from the article by 
Bertini, Freixas, Gambarelli and Stach [@bertini2013comparing]:

```{r, collapse=TRUE, eval=FALSE}
library(CoopGame)
v<-weightedVotingGameVector(n=10, w=c(9,8,7,0,1,1,1,1,1,1), q=23)
normalizedBanzhafIndex(v)
## [1] 0.326633166 0.326633166 0.316582915 0.000000000 0.005025126
## [6] 0.005025126 0.005025126 0.005025126 0.005025126 0.005025126
deeganPackelIndex(v)
## [1] 0.2291667 0.2291667 0.1666667 0.0000000 0.0625000 0.0625000
## [7] 0.0625000 0.0625000 0.0625000 0.0625000
publicGoodIndex(v)
## [1] 0.18181818 0.18181818 0.09090909 0.00000000 0.09090909
## [6] 0.09090909 0.09090909 0.09090909 0.09090909 0.09090909
johnstonIndex(v)
## [1] 0.332692308 0.332692308 0.323076923 0.000000000 0.001923077
## [6] 0.001923077 0.001923077 0.001923077 0.001923077 0.001923077
publicHelpIndex(v)
## [1] 0.15312132 0.15312132 0.15076561 0.07656066 0.07773852
## [6] 0.07773852 0.07773852 0.07773852 0.07773852 0.07773852
publicHelpChiIndex(v)
## [1] 0.16914603 0.16914603 0.16780487 0.06991541 0.07066461
## [6] 0.07066461 0.07066461 0.07066461 0.07066461 0.07066461
#
# Now player 1 donates one vote to player 4
v<-weightedVotingGameVector(n=10, w=c(8,8,7,1,1,1,1,1,1,1), q=23)
normalizedBanzhafIndex(v)
## [1] 0.32908163 0.32908163 0.32397959 0.00255102 0.00255102
## [6] 0.00255102 0.00255102 0.00255102 0.00255102 0.00255102
deeganPackelIndex(v)
## [1] 0.22222222 0.22222222 0.16666667 0.05555556 0.05555556
## [6] 0.05555556 0.05555556 0.05555556 0.05555556 0.05555556
publicGoodIndex(v)
## [1] 0.16666667 0.16666667 0.08333333 0.08333333 0.08333333
## [6] 0.08333333 0.08333333 0.08333333 0.08333333 0.08333333
johnstonIndex(v)
## [1] 0.3329026701 0.3329026701 0.3281653747 0.0008613264 0.0008613264
## [6] 0.0008613264 0.0008613264 0.0008613264 0.0008613264 0.0008613264
publicHelpIndex(v)
## [1] 0.15338882 0.15338882 0.15219976 0.07728894 0.07728894 0.07728894
## [7] 0.07728894 0.07728894 0.07728894 0.07728894
publicHelpChiIndex(v)
## [1] 0.16941222 0.16941222 0.16881669 0.07033698 0.07033698 0.07033698
## [7] 0.07033698 0.07033698 0.07033698 0.07033698
```


### Voting power in the European Parliament (2004 -- 2009)

The article by Aguirre and Quintas [@aguirre2005computing] 
provides an example of voting power in the 
European Parliament as of the election of June 2004.
In the following analysis the seats are allocated to 
eight supranational political groups.
There were 732 members in that parliament, i.e. a majority 
of 367 votes was needed in order to pass a law. 
According to Aguirre and Quintas [@aguirre2005computing] 
the European Democrats (EPP-ED) had 277 members, 
the Party of European Socialists (PES) 198 members,
the European Liberal Demcrat and Reform Party (ELDR) 68 members,
the European Greens / European Free Alliance 
(Greens/EFA) 40 members, 
the European United Left / Nordic Green Left (EUL/NGL) 
39 members,
the Union for a Europe of Nations (UEN) 27 members,
Europe of Democracies and Diversities (EDD) 15 members, 
whereas the rest of 68 parliamentary members 
were not part of any of the seven aforementioned 
supranational political groups.

In the following we confirm the results from table 3 
of the paper by Aguirre and Quintas [@aguirre2005computing]:


```{r, collapse=TRUE, eval=FALSE}
library(CoopGame)
v<-weightedVotingGameVector(n=8, w=c(277,198,68,40,39,27,15,68), q=367)
shapleyShubikIndex(v)
## [1] 0.426190476 0.178571429 0.111904762 0.059523810 0.059523810
## [6] 0.045238095 0.007142857 0.111904762
normalizedBanzhafIndex(v)
## [1] 0.427272727 0.154545455 0.118181818 0.063636364 0.063636364
## [6] 0.045454545 0.009090909 0.118181818
johnstonIndex(v)
## [1] 0.621621622 0.129129129 0.077477477 0.033483483 0.033483483
## [6] 0.023273273 0.004054054 0.077477477
deeganPackelIndex(v)
## [1] 0.2222222 0.1066667 0.1488889 0.1211111 0.1211111 0.1011111
## [7] 0.0300000 0.1488889
publicGoodIndex(v)
## [1] 0.18518519 0.11111111 0.14814815 0.12962963 0.12962963 0.11111111
## [7] 0.03703704 0.14814815
```