---
title: "Introduction to Cooperative Games"
author: "Jochen Staudacher"
date: "6 March 2019"
output: pdf_document
---

# Introduction to Cooperative Games

Today, the term 
game theory is mostly interpreted as 
interactive decision theory,
i.e. game theorists wish to 
study how agents take strategic 
decisions interactively. Since the publication 
of the seminal 
book "Theory of Games and Economic Behavior" by 
John von Neumann and 
Oskar Morgenstern [@NeumannMorgenstern] 
in 1944, modern game theory developed into two main branches. 
These two branches are most frequently referred to as 
noncooperative game theory and cooperative game theory, 
respectively. In our understanding of the field we side with 
the views expressed by Nobel laureate Robert Aumann in his 
famous interview [@vanDammeAumann] with Eric van Damme that 
a better name for noncooperative game theory 
would be "strategically oriented game theory" whereas 
cooperative game might be characterized more precisely 
as "coalitional game theory" or "outcome oriented game theory".

In the introduction of this vignette we wish to borrow 
a thought from the first chapter of the book by 
Peleg and Sudhölter [@PelegSudhoelter]. In the seminal 
papers [@Peleg1996] and [@Peleg1997] Bezalel 
Peleg pointed out 
that Nash's program could not be implemented, i.e. that 
it is not possible to express each and every cooperative 
game as a noncooperative game in extensive
form with the solution of the cooperative game
being defined in terms of equilibrium points of the 
corresponding noncooperative game. 

In other words: Peleg's results from 
[@Peleg1996] and [@Peleg1997] removed any doubts 
that cooperative game theory was truly a theory 
of its own right.
So, to quote Robert Aumann (see [@vanDammeAumann], p. 195) again 
```
... It is not only strategic interaction.
```
In this spirit our R-package `CoopGame` is devoted 
to the study of coalitional games 
with transferable utility and aims to provide a 
comprehensive set of methods. 

## What this package provides and what it does not provide

The package `CoopGame` focusses on a cost-savings approach to
cooperative games.

We are studying a transferable utility game (TU game) $v$ in 
characteristic function form consisting of the 
player set $N =\{ 1, \dots , n \}$ and the characteristic
function $v: 2^{N} \to \mathbb{R}$ with $v (\emptyset) = 0$.
We specify such a TU game with $n$ players as a (game) vector 
of length $2^n - 1$.

### Vignette for version 0.2.1 of `CoopGame`

This is the place to emphasize once again that this vignette 
accompanies version 0.2.1 of the `R` package `CoopGame`. 
Future releases of `CoopGame` might incorporate additional 
functionality and will be accompanied by an update of this 
vignette. 

### Specifying a TU game

Let us look at an introductory example of a three-player 
TU game. Let $N= \{ 1, 2, 3 \}$ and 
$v: \mathcal{P}(N) \to \mathbb{R}$ 
be specified as   
$$ 
\begin{aligned}
v(\emptyset) & =  0, \\  
v(\{ 1 \}) & =  0, \\
v(\{ 2 \}) & =  0, \\ 
v(\{ 3 \}) & =  0, \\
v(\{ 1,2 \}) & =  100, \\ 
v(\{ 1,3 \}) & =  115, \\ 
v(\{ 2,3 \}) & =  125, \\
v(\{ 1,2,3 \}) & =  220. 
\end{aligned}
$$
We may interpret the values of the individual coalitions 
in various ways, e.g. as the worth of a coalition or the 
cost-savings of a coalition. In our above example the singleton 
coalitions have worth zero whereas the grand coalition 
$N= \{ 1, 2, 3 \}$ makes a total of $220$ (monetary units). 
In brief, the 
basic question in coalitional games with transferable utility 
is how we can share the worth of $220$ among our three players.
In `R` we can simply specify the above game as a vector $v$ of 
lenght $7$, i.e.
```{r, collapse=TRUE}
library(CoopGame)
(v <- c(0,0,0,100,115,125,220))
```

Sometimes it is convenient to use the shorthand notations 
$$ v_{i} = v(\{ i \}) \quad \textrm{for} \quad i=1, \dots , n, $$
for the singleton coalitions.

### A popular example: The Maschler game

Throughout this vignette we will frequently employ the 
following three-player example
$$ 
v(C) = \Bigg\{  
\begin{aligned}
0, & \quad \vert C \vert = 1 \\
60, & \quad \vert C \vert = 2 \\
72, & \quad \vert C \vert = 3
\end{aligned}
$$

It appears in the article [@AumannMyerson] where the 
authors attribute it to the famous Israeli 
game theorist Michael Maschler. The above 
game has thus frequently been 
referred to as the Maschler game in the literature since.

```{r, collapse=TRUE}
library(CoopGame)
(Maschler <- c(0,0,0,60,60,60,72))
```

### Functionality and structure of the package `CoopGame`

The package `CoopGame` provides functions for 

* generating TU games with special structure (see chapter 2 
of this vignette), like 
e.g. bankruptcy games, cost games, weighted voting 
games and unanimity games 
* checking game properties (see chapter 3 of this vignette), 
like e.g. superadditivity, convexity and balancedness
* computing a number of set-valued solution concepts for 
TU games (see chapter 4 
of this vignette), including the core and various core catchers
* computing a large array of point-valued solution concepts 
for TU games (see chapter 5 
of this vignette), 
including the Shapley value, the nucleolus as well as 
various nucleolus derivatives and numerous power indices 
* drawing both set- and point-valued solution concepts 
for the 3- and 4-player cases (see chapter 6 of this vignette)
* some general functionality useful in the context of TU games 
(see section 1.2 of this vignette), like e.g. computing the 
unanimity coefficients for a given game vector

### Models and questions beyond the scope of `CoopGame`

Coalitional games can be turned into more realistic 
models if, in addition one allows for partitions of the player 
set (see e.g. the original paper by Thrall and Lucas [@thrall1963n])
or for specifying an undirected graph connecting the players. 
The latter case is frequently referred to as a communication 
game (see e.g. the book by 
Slikker and van den Nouweland [@SlikkerNouweland]). The 
authors are currently 
developing additional `R` packages for both games 
with partitions 
of the player set as well as communication games. These 
additional software packages will make use of `CoopGame`, 
but `CoopGame` itself will allow for neither partitions or 
communication structures. 

Since `CoopGame` can neither handle partitions nor 
other more general coalition structures (see e.g. the 
paper by Aumann and Dreze [@aumann1974cooperative] from 1974), the 
package neither provides methods for computing 
bargaining sets (see e.g. the original paper by 
Aumann and Maschler [@aumann1964bargaining] from 1964) nor
methods for computing the 
kernel (see e.g. the original article by Aumann, Peleg and 
Rabinowitz [@aumann1965method] from 1965). 

Another feature this first release of `CoopGame` 
also lacks is functionality for general market 
games (see e.g. the original paper by Aumann [@aumann1964markets] 
on markets with a continuum of traders from 1964).

Also, NTU games, i.e. cooperative games with non-transferable 
utility (see e.g. the article by Aumann and 
Maschler [@AumannMaschlerNTU] from 1960), 
are beyond the scope of 
this `R` package.

### This is not a book on cooperative game theory and these are not lecture notes 

This text is a CRAN vignette only and it is supposed to 
give users of `CoopGame` a concise overview of the 
package. Users looking for an introduction to the 
fascinating field of cooperative game theory will be 
disappointed by this text. We strictly adhere to the CRAN 
policy that CRAN vignettes are not to be misused as 
lecture notes. 

For excellent introductions to cooperative 
game theory we refer to the books by 
Peleg and Sudhölter [@PelegSudhoelter], 
by Driessen [@DriessenBook], 
by Branzei, Dimitrov and Tijs [@BranzeiBook], 
by Chakravarty, Mitra and Sarkar [@ChakravartyMitraSarkar] 
and by Gilles [@GillesBook].

Excellent introductions to cooperative game theory 
can also be found in the game theory books 
by Maschler, Solan and Zamir
[@MaschlerSolanZamir], by Peters [@PetersBook]
by Osborne and Rubinstein [@OsborneRubinsteinBook], 
the recent book by Narahari [@NarahariBook] and the 
classic textbook by Straffin [@StraffinBook].

We finally wish to acknowledge that beyond the aforementioned 
excellent books 
the development of 
`CoopGame` also benefitted greatly from two 
excellent German textbooks, i.e. the books by 
Wiese [@wiese2005kooperative] and by 
Holler and Illing [@HollerIlling].


## Some general functionality for cooperative games 

In this section we introduce some useful functionality 
for cooperative games which we will use later when we 
discuss games with special structure, game properties as well as 
set-valued and point-valued solution concepts for cooperative 
games.

### Zero-normalized and zero-one-normalized games

We call a cooperative game $v$ **zero-normalized** if 
$$ v_{i} = 0 \quad \textrm{for} \quad i=1, \dots , n, $$
i.e. the values of all singleton coalitions are zero 
(see e.g. Peleg and Sudhölter [@PelegSudhoelter], p. 11).

We can easily zero-normalize a given game vector $v$ into a 
corresponding zero-normalized game vector $w$ via 
$$ w (C) = v(C) - \sum_{i \in C} v_{i} $$
for every coalition $C \in \mathcal{P}(N)$ 
(see e.g. Branzei, 
Dimitrov and Tijs [@BranzeiBook], p. 9).
In `CoopGame` we provide a corresponding 
function `getZeroNormalizedGameVector`:
```{r, collapse=TRUE}
library(CoopGame)
v <- c(30,40,50,90,100,110,180)
(w <- getZeroNormalizedGameVector(v))
```

We call a cooperative game $v$ **zero-one-normalized** if 
it is zero-normalized and for the grand coalition $N$ there 
holds $v(N) = 1$. 
We can easily zero-one-normalize a given game vector $v$ by 
dividing the zero-normalized game vector $w$ by the value 
of the grand coalition $v(N)$.
In `CoopGame` we provide a corresponding 
function `getZeroOneNormalizedGameVector`:
```{r, collapse=TRUE}
library(CoopGame)
v <- c(30,40,50,90,100,110,180)
(w01 <- getZeroOneNormalizedGameVector(v))
```

For a detailed discussion on the importance 
of zero-one-normalized 
games and strategic equivalence we refer the reader to 
the book by Maschler, 
Solan and Zamir [@MaschlerSolanZamir], p. 670.

### Why bit matrices are helpful in cooperative game theory

We introduce the concept of a bit matrix as we find it 
extremely useful for working with cooperative games. 
Bit matrices unambigously map the values of coalitions 
to the players in that coalition. We again make use of 
the Maschler game for our example.

```{r, collapse=TRUE}
library(CoopGame)
(Maschler <- c(0,0,0,60,60,60,72))
createBitMatrix(n=3,Maschler)
```

For a TU game with $n$ players the function `createBitMatrix` 
creates a bit matrix with $n+1$ columns and 
$2^n-1$ rows which contains all possible 
coalitions of the players (apart from the null coalition).
Each player is represented by a column which states 
if this player is participating in a coalition (value $1$) 
or not (value $0$). The last 
column (named `cVal`) contains 
the values of each coalition. Accordingly, each row of the 
bit matrix expresses a coalition as a subset of all players.

However, we need to admit that our usage of bit 
matrices in `CoopGame` also consumes plenty of storage space. 

### Marginal contributions

Let our TU game be specified by its characteristic function $v$.
Then for every player $j \in N$ and for every coalition 
$C \in \mathcal{P}(N)$ with $j \in C$ we can define the 
so-called **marginal contribution** of player $j$ to coalition $C$ 
as
$$ v(C) - v(C \backslash \{ j \}), $$
see e.g. the book by Branzei, Dimitrov and Tijs [@BranzeiBook], pp. 6, 
or the book by Peters [@PetersBook], pp. 156. 
For some point- and set-valued solution concepts we need to 
know the marginal contributions of every player for every 
permutation of the set of players.
Our function `getMarginalContributions` provides the user 
with a list of all combinations, i.e. permutations of 
the players, used and 
a corresponding matrix of marginal contributions:
```{r, collapse=TRUE}
library(CoopGame)
v <- c(3,4,5,9,10,11,18)
(MC <- getMarginalContributions(v))
# Look at all the permutations computed
MC$combinations
# Look at the matrix of marginal values 
# corresponding to these permutations
MC$marginal_values
```
It appears worthwhile to interpret the second line of 
the marginal values we just computed 
```{r, collapse=TRUE}
MC$marginal_values[2,]
```
along the lines of the book by Peters [@PetersBook], pp. 156.
The above results corresponds to the permutation $(1,3,2)$. 

- Player $1$ enters first and contributes $3$. 
- Player $3$ enters second and contributes $11-3=8$. 
- Player $2$ enters last and contributes $18-11=7$. 

### The dual game 
Let our TU game be specified by its characteristic function $v$.
Then we can specify the so-called **dual game** $v^{*}$ 
corresponding to $v$ via 
$$ v^{*} (C) = v(N) - v(N \backslash C) $$
for every coalition $C \in \mathcal{P}(N)$ (see 
e.g. the book by Peleg and Sudhölter [@PelegSudhoelter], p. 125, 
or the book by Branzei, Dimitrov and Tijs [@BranzeiBook], p.7, 
for more details). 
The package `CoopGame` provides a function `getDualGameVector`:
```{r, collapse=TRUE}
library(CoopGame)
v <- c(3,4,5,9,10,11,18)
# Compute dual game vector
(vStar <- getDualGameVector(v))
```

### The utopia payoff vector 
In cooperative game theory the so-called 
**utopia payoff** of player $j$ is defined as 
$$ M_{j} = v(N) - v(N \backslash \{ j \}) \quad \textrm{for} \quad j=1, \dots , n,$$
i.e. the utopia payoff $M_{j}$ is the 
marginal contribution of player $j$ to the grand 
coalition $N$ (see e.g. the book by 
Branzei, Dimitrov and Tijs [@BranzeiBook], p. 20). 
The package `CoopGame` provides a function 
`getUtopiaPayoff` for computing a vector of 
utopia payoffs for all players:
```{r, collapse=TRUE}
library(CoopGame)
v <- c(3,4,5,9,10,11,18)
# Compute utopia payoff vector for specified game v
(M <- getUtopiaPayoff(v))
```
It is clear that player $j$ can not ask for more 
than $M_{j}$ in the grand coalition.
The utopia payoff vector 
will lateron play a role in defining game properties 
as well as point- and set-valued solution concepts. 

### The minimal rights vector 
The so-called remainder $R(C, j)$
of player $j$ in coalition $C \in \mathcal{P}(N)$ is 
the amount which remains for player $j$ if the 
coalition $C$ forms and the rest of the players in 
coalition $C$ all obtain their 
individual utopia payoffs, i.e.
$$ R(C,j) = v(C) - \sum_{k \in C, k \neq j} M_{k}. $$ 
We can define a vector of **minimal rights** with components 
$$ m_{j} = \max_{C: j \in C} R(C,j),  \quad \textrm{for} 
\quad j=1, \dots , n,$$
since player $j$ has a justification to demand at 
least $m_{j}$ in the grand coalition, 
see e.g. the book by 
Branzei, Dimitrov and Tijs [@BranzeiBook], p. 20.  
The package `CoopGame` provides a function 
`getMinimalRights` for computing a vector of 
minimal rights for every player:
```{r, collapse=TRUE}
library(CoopGame)
v <- c(2,3,4,8,9,10,13)
# Compute minimal rights vector for specified game v
(m <- getMinimalRights(v))
```

### The excess coefficients

The excess $e(C,x)$ of a coalition $C$ with respect to a vector 
$x$ measures the gain or loss of the members of $C$ in case they 
decide to abandon the grand coalition $N$ in favour of their 
own coalition $C$, see e.g. the book by 
Driessen [@DriessenBook], p. 12:

$$ e(C,x) = v(C) - \sum_{j \in C} x_{j} = v(C) - x(C)$$ 
In the above formula we use the 
shorthand notation $x(C) =  \sum_{j \in C} x_{j}$.

The package `CoopGame` provides a function 
`getExcessCoefficients` for computing a vector of 
excess coefficients for every coalition:
```{r, collapse=TRUE}
library(CoopGame)
A <-c(3,4,5,9,10,11,18)
x <-c(5,6,7)
# Compute vector of excess coefficients for specified game v
(ec <- getExcessCoefficients(A,x))
```
Note that the last component of a vector of excess coefficients 
is always $0$ as long as $x$ is efficient, i.e. 
$\sum_{j \in C} x_{j} = v(N)$.

The concept of excesses is important in various 
solution concepts, like e.g. the 
nucleolus (see e.g. the original paper 
by Schmeidler [@schmeidler1969nucleolus] which appeared in 1969). 
Computing a vector of excesses comes in handy for checking 
the correctness of nucleolus or prenucleolus computations, 
see the article by Guajardo and 
Jörnsten [@CommonMistakesNucleolus].

### The gap function
The gap function is the additive inverse of the vector of 
excesses with respect to the utopia payoff vector, see the 
book by Driessen [@DriessenBook], p. 57, for more details. 

The package `CoopGame` provides a function 
`getGapFunctionCoefficients` for computing the vector of 
gap function coefficients for every coalition:
```{r, collapse=TRUE}
library(CoopGame)
A <-c(3,4,5,9,10,11,18)
# Compute vector of gap function coefficients for specified game v
(gap <- getGapFunctionCoefficients(A))
```


### The per capita excess coefficients
Per capita excess coefficients replace excess coefficients 
in the computation of the 
per capita 
nucleolus (see the original paper by Young [@Young1985]).
The function 
`getPerCapitaExcessCoefficients` computes a vector of 
per capita excess coefficients for every coalition 
and a vector $x$, 
i.e. the excess coefficients are divided by the number 
of players in each coalition:
```{r, collapse=TRUE}
library(CoopGame)
A <-c(3,4,5,9,10,11,18)
x <-c(5,6,7)
# Compute vector of per capita excess coefficients for specified game v
(ecpc <- getPerCapitaExcessCoefficients(A,x))
```

Computing a vector of per capita excess coefficients 
comes in handy for checking 
the correctness of computations
of the per capita nucleolus, see [@CommonMistakesNucleolus].

### Propensities to disrupt
For a cooperative game $v$ and a payoff 
vector $x$ with $\sum_{j=1}^{n} x_{j} = v(N)$
player $i$'s propensity to disrupt (see 
e.g. the article by Littlechild 
and Vaidya [@LittlechildVaidya]) is defined as  
$$
d(i,x) = \frac{\sum_{j=1,j \neq i}^{n} x_{j} - v(N \backslash \{ i \})}{x_{i} - v_{i}}
$$
The above expression quantifies the disruption caused 
if player $i$ breaks away from the grand coalition.
Within this expression the 
denominator stands for the loss incurred by 
player $i$ for breaking away from the grand coalition, 
whereas the numerator stands for the joint loss of the 
rest of the players due to the breakup caused by player $i$.
This concept is important in solution concepts like the 
Gately point (see e.g. the original paper by 
Gately [@GatelyPaper] from 1974, the paper by 
Littlechild and Vaidya [@LittlechildVaidya] or the 
recent article [@StaudacherAnwander] by the authors) and 
the disruption nucleolus (see [@LittlechildVaidya]). 

The function 
`getVectorOfPropensitiesToDisrupt` computes a vector of 
propensities to disrupt for every coalition 
and a vector $x$:
```{r, collapse=TRUE}
library(CoopGame)
A <-c(3,4,5,9,10,11,18)
x <-c(5,6,7)
# Compute vector of propensities to disrupt for specified game v
(propVec <- getVectorOfPropensitiesToDisrupt(A,x))
```
Note that the last component of a vector of propensities 
to disrupt 
is always set to $0$.
Computing a vector of coefficients 
of propensities to disrupt 
comes in handy for checking 
the correctness of computations
of the disruption nucleolus, see [@CommonMistakesNucleolus].

### The equal propensity to disrupt 
`CoopGame` also provides a function for computing the 
so-called equal propensity to disrupt. 
This concept originates from 
the paper [@GatelyPaper] by Gately for 
three-person games and was generalized to $n$-person games 
by Littlechild and 
Vaidya (see [@LittlechildVaidya], p. 152). The 
goal is to find an imputation $x$ 
with minimal propensity to disrupt. 
It can be shown that this minimal propensity to 
disrupt can be found 
by equating the propensity to disrupt over all players, i.e.
$$ d(i,x) = d^{*} \quad \textrm{for} \quad i=1, \dots , n.$$
As pointed out by 
Littlechild and Vaidya (see [@LittlechildVaidya], p. 153) using 
the above approach one can easily 
find the following closed-form expression 
$$d^{*} = \frac{(n-1) v(N) - \sum_{j=1}^{n} v(N \backslash \{ j \})}{v(N) - \sum_{j=1}^{n} v_{j}} = \frac{\sum_{j=1}^{n} M_{j} - v(N)}{v(N) - \sum_{j=1}^{n} v_{j}}$$
for the equal propensity to disrupt $d^{*}$ of 
a TU game with $n$ players. Our recent paper 
[@StaudacherAnwander] provides some insight and interpretation 
for the cases $d^{*} = 0$ and $d^{*} < 0$.


### Minimum winning coalitions and real gaining coalitions
In a simple game $v$ we call a player $j$ critical for 
a coalition $C$ if the departure of player $j$ turns 
$C$ from a winning coalition into a losing coalition, 
i.e. $v(C)=1$ and $v(C \backslash j)=0$.
A minimum winning coalition in a simple game is a coalition 
where every member of the coalition is critical, see e.g. 
the paper by Deegan and Packel [@DeeganPackel], 
the paper by Holler [@Holler1982] or 
the article by Bertini [@bertini2011minimal].
The function `getMinimumWinningCoalitions` 
identifies all minimal winning coalitions
in a simple game $v$ and returns a corresponding 
data frame. The function
`getCriticalCoalitionsOfPlayer` 
identifies all coalitions for which a given player 
is critical.
```{r, collapse=TRUE}
library(CoopGame)
# Define a simple game
A <-c(0,0,0,1,1,0,1)
# Find the minimum winning coalitions
getMinimumWinningCoalitions(A)
# Find the coalitions where player 2 is critical
getCriticalCoalitionsOfPlayer(2,A)
```
The concept of minimum winning coalitions 
can be generalized to general cooperative 
games $v$ via the concept of real gaining 
coalitions (see e.g. the original 
paper by Holler and Li [@holler1995public] or 
the article by Bertini and Stach [@BertiniStach2015]). 
$C$ is called
a real gaining coalition (RGC) iff
for any true subset $T \subset C$ there holds $v(C)-v(T) > 0$. 
We provide a corresponding 
function `getRealGainingCoalitions`.
```{r, collapse=TRUE}
library(CoopGame)
A <-c(0,0,0,0.8,0.9,0,0.9)
# Find the real gaining coalitions
getRealGainingCoalitions(A)
```

We provide corresponding functions `getWinningCoalitions` 
and `getGainingCoalitions` for computing the winning 
coalitions in simple games (see e.g. the 
original paper by Bertini, Gambarelli 
and Stach [@BertiniGambarelliStach])
and the gaining coalitions in general 
TU games (see e.g. the original paper 
by Bertini and Stach [@BertiniStach2015] from 2015), 
respectively. 

### The unanimity coefficients
The unanimity coefficients represent a 
cooperative game in an alternative basis, 
the so-called unanimity basis. 
They were introduced in the seminal paper 
by Shapley [@shapley1953value] 
and are also called Harsanyi dividends (see also the books 
by Peleg and Sudhölter [@PelegSudhoelter], p. 153, 
or by Gilles [@GillesBook], pp. 15--17, for more details). 

Our package provides a function `getUnanimityCoefficients`.

```{r, collapse=TRUE}
library(CoopGame)
# The Maschler game
Maschler <-c(0,0,0,60,60,60,72)
# Find the unanimity coefficients for the Maschler game
(unCoeff <- getUnanimityCoefficients(Maschler))
```

### The k-cover

In [@DriessenBook], p. 173, Driessen defines an associated cover 
$v_{k}$ that majorizes the original game $v$ for a given integer $k$.
$$
v_{k}(C) = \left\{
\begin{array}{rl}
v(C) & \quad \text{if } \quad \vert C \vert < k,\\
v(N) - \sum_{j \in N \backslash C} M_{j} & \quad \text{else,}
\end{array} \right. 
$$
with $M$ denoting the utopia payoff vector. 
In case the gap function $g$ of original game $v$ satisfies 
$$0 \leq g(N) \leq g(C)$$
for all coalitions $C \subseteq N$ with $k$ or more elements,
then $v_{k}$ is called 
the k-cover of the game $v$. In `CoopGame` 
we provide a function 
`getkCover` computing $v_{k}$ according to 
the previous formula. 
In case $v$ does satisfy the above condition, `getkCover` 
will return `NULL`. 

```{r, collapse=TRUE}
library(CoopGame)
# Example from book by T. Driessen, p. 75
A=c(0,0,0,9,9,15,18)
# Compute 1-cover of this 1-convex game
(A1 <- getkCover(A,k=1))
```

## Cooperative game theory in the `R` Ecosystem
There already exist two small `R` packages offering 
functionality for 
cooperative game theory on CRAN, i.e. 
`GameTheoryAllocation` [@GameTheoryAllocation] and 
`GameTheory` [@cano2017enjoying]. Both 
packages are very limited in scope, e.g. according 
to its documentation [@cano2017enjoying] the package 
`GameTheory` only computes the nucleolus for a maximum 
of four players. However, `GameTheory` provides some 
very nice insight into conflicting claims problems 
and is also employed in the `R` 
package `coopProductGame` [@coopProductGame] 
for computing linear production 
games (see the original paper by 
Owen [@owen1975core] from 
1975 for more details on 
linear production games). 

Still, we found it very sensible to develop `CoopGame`. 
Our goal was to provide a comprehensive package 
for cooperative game theory that goes beyond providing 
routines for the Shapley value or the nucleolus. 
One of our goals was to provide reference implementations 
for lesser-known and lesser-used solution concepts.
Another was to be able to visualize solution concepts 
in the case of three or four players.

## A few general remarks on `CoopGame`

We wish to end this first chapter of our vignette with 
four general remarks on `CoopGame`.

### Package maintainer 

This is supposed to be the second `R` package on CRAN 
maintained by the first author after the publication 
of the significantly smaller package `EvolutionaryGames` 
[@EvolutionaryGames] in November 2017. Please feel free 
to email questions regarding `CoopGame` to 
`jochen.staudacher@hs-kempten.de`.

### No formal maximum number of players
`CoopGame` does not enforce a maximum number of players. 
Still, realistically for most concepts provided in 
`CoopGame` users should limit their studies to 
at most $20$ players if they expect quick answers 
on modern computers. The authors have employed `CoopGame` for 
up to $n=24$ players, but `R` normally ran out of storage 
for $n>24$.

### Minimum: 2 Players
In `CoopGame` a minimum of two players is needed 
in order to define a TU game.

### No null games allowed
In case the user specifies a null game, i.e. a trivial TU game 
with each coalition generating zero value, `CoopGame` 
will stop with an error message.
