---
title: "VisualizationFor3And4PlayerGames"
author: "Jochen Staudacher"
date: "6 March 2019"
output: pdf_document
---

# Visualization of Solution Concepts for Games with 3 and 4 Players

The package `CoopGame` offers the possibility to 
visualize both set valued and point valued solution 
concepts for cooperative games with 3 or 4 players 
using barycentric coordinates. 
We strongly hope that 
our visualization routines will be helpful for 
colleagues teaching cooperative game theory and 
wish to acknowledge that for the conversion of 
Cartesian coordinates to barycentric coordinates 
`CoopGame` makes use of the package `geometry` [@geometry].

## An example: Generating the `CoopGame` Logo

Running the following lines of `R` Code users can generate 
the `CoopGame` Logo on the first page of this vignette. 
It draws the reasonable set of the given game (in blue) as 
a subset of 
the imputation set and the core of the game (in red) as 
a subset of the reasonable set. The modiclus is displayed 
in black as a point in the core.

```{r, collapse=TRUE, fig.show=TRUE}
library(CoopGame)
v0=c(6,8,10,18,20,22,31)
drawImputationset(v0, label=FALSE)
drawReasonableSet(v0, colour="blue", holdOn=TRUE)
drawCore(v0, holdOn=TRUE, colour="red")
drawModiclus(v0, holdOn=TRUE, colour="black")
```

## Some general remarks on the drawing routines in `CoopGame`
The package `CoopGame` provides drawing routines for all 
set solution concepts from chapter 4 for the case of 
3 or 4 players.
In particular, there are routines `drawImputationSet`, 
`drawCore`, `drawCoreCover`, 
`drawReasonableSet` and `drawWeberset`.
Note that core, core cover, reasonable set and Weber set 
are all visualized as parts of the imputation set. 
Note also that the Weber set is only guaranteed to be a subset 
of the imputation set for superadditive games 
whereas the other three set solution 
concepts are always subsets of the imputation set.
Also, any efficient point solution method, i.e. any 
point solution method for which the components of the 
solution vector are guaranteed to add up to the value 
of the grand coalition, come with their own drawing routines.

As for the individual drawing routines users can add to 
an existing plot by setting the parameter `holdOn` to 
`TRUE`. The parameter `holdOn` is set to `FALSE` by 
default and this feature should be handled with a little care 
by the user. We strongly advise against plots with too 
many set valued and/or point valued solutions.

In addition, users can also set the `colour` of the 
polyhedron or point they want to draw and use `name` 
for naming an individual point. Only for the function 
`drawImputationset` the colour is fixed to white as 
any polyhedra or points are always displayed as 
parts of the imputation set. 

Also, only for the function 
`drawImputationset` there is the default `label=TRUE`.
In any plots the vertices of the imputation set are 
by default 
labelled by their coordinates. In other words: 
Unless users start 
with `drawImputationset(gameVector, label=FALSE)` 
and then carefully add to their 
plots with `holdOn=TRUE` the 
points of the imputation set will by default be labelled. 
For all four drawing 
routines `drawCore`, `drawCoreCover`, 
`drawReasonableSet` and `drawWeberset` the parameter 
`label` is set to `FALSE` by default, i.e. by default 
the coordinates of the vertices of these sets 
are not displayed. The following code yields a 
variant of the `CoopGame` 
logo with the three corner points of 
the imputation set labelled. 

```{r, collapse=TRUE, fig.show=TRUE}
library(CoopGame)
v0=c(6,8,10,18,20,22,31)
drawReasonableSet(v0, colour="blue")
drawCore(v0, holdOn=TRUE, colour="red")
drawModiclus(v0, holdOn=TRUE, colour="black")
```

Placing labels 
in the perfect places by default is tricky 
and users may not always be perfectly satisfied 
with the results they obtain when plotting set solution 
concepts with `label = TRUE`. So our advice 
to users intending to use `CoopGame` for publication ready 
graphics of the core, the core cover, the reasonable 
set and/or the Weber set 
with all points labelled by their coordinates 
is to export the plots without these 
labels and then postprocess the plots themselves putting 
labels of individual points in the desired places.

For drawing any point valued solution concepts `label` 
is actived by default and a proper `name` for the point 
valued solution is set by default.

## The power of `rgl`

The R package `rgl` [@rgl] offers the possibility to 
visualize solution concepts for cooperative games with 
4 players in three dimensions. Users are invited to 
uncomment the following four lines of code,
play around with the following example and rotate the 
nice 3d-plot displayed by `rgl`.

```{r, collapse=TRUE, fig.show=FALSE, eval=FALSE}
#library(CoopGame)
#A=c(0,0,0,0,8,8,8,8,8,8,12,12,12,12,22)
#drawWeberset(A, colour ="grey")
#drawCore(A, colour ="red", holdOn = TRUE)
```