---
title: "Game Properties"
author: "Jochen Staudacher"
date: "6 March 2019"
output: pdf_document
---


# Game Properties

The package `CoopGame` provides 
users with the possibility to check 
a given game vector $v$ for a number of different game properties.
In this chapter of our vignette we would like to give 
a quick overview of these game properties and the 
precise definitions we use. 

## Checking a game property 

Given that the $17$ different functions for 
checking various game properties all come with 
detailed examples in their 
individual documentations we will only specify 
the basic principle for checking a game property. 
Users simply need to specify a game vector and then check 
the corresponding game property (which always starts with `is`).
The result will be `TRUE` if 
the game shares the property in question, else it will be `FALSE`.

```{r, collapse=TRUE}
library(CoopGame)
A <-c(0,0,0,1,1,0,1)
isSuperadditiveGame(A)
```

## A quick overview of available game properties

### Nonnegative games 

We call a game vector $A$ **nonnegative** if all of its entries 
are nonnegative. This game property can be checked via the 
function `isNonnegativeGame`. We provide certain 
solution concepts 
for nonnegative games only.

### Essential games

We call a TU game $v$ **essential**, if the value of the 
grand coalition $v(N)$ is greater than the 
sum of the values of the singleton 
coalitions $v(\{i\})$, i.e.

$$v(N) > \sum_{i=1}^{n} v(\{i\}).$$

Our definition follows the books by Chakravarty,
Mitra and Sarkar [@ChakravartyMitraSarkar], p. 23, 
and by Gilles [@GillesBook], p. 18.
We find it very convenient that according to this definition 
the imputation set of an essential game is nonempty and consists 
of more than one point.
This game property can be checked via the 
function `isEssentialGame`.

### Degenerate games

We call a TU game $v$ **degenerate** (or **inessential**), if 
the value of the 
grand coalition $v(N)$ equals the 
sum of the values of the singleton 
coalitions $v(\{i\})$, i.e.

$$v(N) = \sum_{i=1}^{n} v(\{i\}).$$
We find it very convenient that according to this definition 
the imputation set of a degenerate game consists
of exactly one point (specified by the singleton coalitions).
This game property can be checked via the 
function `isDegenerateGame`.

### Monotonic games

According to the book by 
Peleg and Sudhölter [@PelegSudhoelter], p. 12, 
we call a TU game $v$ **monotonic** if 

$$ S \subseteq T \subseteq N \Rightarrow v(S) \leq v(T). $$
This game property can be checked via the 
function `isMonotonicGame`.

### Simple games

Along the lines of the book by 
Peleg and Sudhölter [@PelegSudhoelter], p. 16, 
we call a TU game $v$ **simple** if 
$v$ is monotonic and the values of all coalitions are 
$0$ or $1$. This game property can be checked via the 
function `isSimpleGame`.

### Symmetric games

Following the book by 
Peleg and Sudhölter [@PelegSudhoelter], p. 12, 
we call a TU game $v$ **symmetric** if
the values of all 
coalitions containing the same number of players are identical, i.e.

$$ \vert S \vert = \vert T \vert \Rightarrow v(S) = v(T) $$
for all coalitions $S,T \subseteq N$. 
This game property can be checked via the 
function `isSymmetricGame`.

### Constant-sum games 

In a **constant-sum** game $v$ for any coalition $S$ the sums
of $v(S)$ and its complement $v(N \backslash S)$ equal
$v(N)$, i.e. 
$$ v(S) + v(N \backslash S) = v(N) $$
for all $S \subseteq N$, see e.g. 
the book by Peleg and Sudhölter 
[@PelegSudhoelter], p. 11.
This game property can be checked via the 
function `isConstantSumGame`.

### Weakly constant-sum games
We call a TU game $v$ with $n$ players
**weakly constant-sum** as long as
the constant-sum condition holds for coalitions of
sizes $1$ and $n-1$, i.e.
$$ v(\{i\}) + v(N \backslash \{i\}) = v(N) $$
for all players $i=1, \dots, n$. In other words:
For weakly constant-sum games the vector of singleton
coalitions and the utopia payoff vector coincide.
This game property comes in handy when checking
uniqueness conditions for the Gately point, see
the paper [@StaudacherAnwander] by the authors 
of this vignette. Also note that
any constant-sum game is weakly constant-sum, but not
vice versa.
This game property can be checked via the 
function `isWeaklyConstantSumGame`.


### Superadditive games

We call a TU game $v$ **superadditive** if 

$$ v(S \cup T) \geq v(S) + v(T) $$
for all coalitions $S, T \subseteq N$ with $S \cap T = \emptyset$,
see e.g. the book by 
Peleg and Sudhölter [@PelegSudhoelter], p. 10, or 
the book by Maschler, Solan and 
Zamir [@MaschlerSolanZamir], p. 671.
The idea of superadditivity is that disjoint groups of players
are never punished for cooperating.
This game property can be checked via the 
function `isSuperadditiveGame`.


### Additive games

We call a TU game $v$ **additive** if 

$$ v(S \cup T) = v(S) + v(T) $$
for all coalitions $S, T \subseteq N$ with $S \cap T = \emptyset$,
see e.g. the 
book by Peleg and Sudhölter [@PelegSudhoelter], p. 11, 
or the book by Maschler, Solan and Zamir 
[@MaschlerSolanZamir], p. 792.
Note that an additive game is always superadditive, constant-sum,
weakly constant-sum and degenerate whereas none of these four
game properties guarantee additivity in return.
This game property can be checked via the 
function `isAdditiveGame`.

### Weakly superadditive games

We call a TU game $v$ with $n$ players **weakly superadditive** if 

$$ v(S \cup \{ i \}) \geq v(S) + v(\{ i \}) $$
for all coalitions $S \subseteq N$ and
all players $i=1, \dots ,n$ with $i \notin S$,
see e.g. the book by 
Peleg and Sudhölter [@PelegSudhoelter], p. 10. Note that
weak superadditivity is equivalent to 0-monotonicity, i.e. the
zero-normalization of the game is monotonic, see
e.g. Maschler, Solan and Zamir [@MaschlerSolanZamir], 
p. 789, or Peleg and Sudhölter [@PelegSudhoelter], p. 12.
This game property can be checked via the 
function `isWeaklySuperadditiveGame`.

### Quasi-balanced games
Given a TU game $v$ with $n$ players let $m$ denote
the minimal rights vector and $M$ the utopia payoff vector.
We call the TU game $v$ **quasi-balanced** if
$m(i) \leq M(i)$ for all $i=1, \dots, n$ and 
$$ \sum_{i=1}^{n} m_{i} \leq v(N) \leq \sum_{i=1}^{n} M_{i}, $$
see e.g. the book 
by Branzei, Dimitrov and Tijs [@BranzeiBook], p. 31.
Quasi-balanced games can equivalently be characterized
as the TU games with a non-empty core cover.
Note also that the $\tau$-value, an important solution
concept, is only defined for quasi-balanced games,
see e.g. the original paper by Tijs [@Tijs1981] 
from 1981.
This game property can be checked via the 
function `isQuasiBalancedGame`.

### Balanced games
A TU game $v$ is called **balanced** if and only if the core
of $v$ is non-empty. For the sake of brevity we
do not go into any details of the balancedness condition
and the Bondareva-Shapley theorem in this vignette.
Instead, we only refer to the original papers by 
Bondareva [@bondareva1963some] and by Shapley
[@shapley1967balanced] as well as to section 3.1 of
the book by Peleg and Sudhölter [@PelegSudhoelter] for
further details.
This game property can be checked via the 
function `isBalancedGame`.

### Convex games 
We call a TU game $v$ **convex** if 

$$ v(S \cup T) + v(S \cap T) \geq v(S) + v(T) $$
for all $S, T \subseteq N$, see e.g. the book by 
Peleg and Sudhölter [@PelegSudhoelter], p. 10, 
for more details. Convex
games are always
balanced, i.e. the core of a convex game is never empty, and
they arise in various important application areas of
cooperative game theory.
This game property can be checked via the 
function `isConvexGame`.

### Semiconvex games
A TU game $v$ with $n$ players is called **semiconvex**
if for its gap function $g$ there holds
$$0 \leq g(i) \leq g(S)$$
for all players $i=1, \dots, n$ and
all coalitions $S \subseteq N$ with $i \in S$.
Note that convex games are always semiconvex but not
vice versa.
We refer to the book by Driessen [@DriessenBook], p. 76, or
the original paper by Driessen 
and Tijs [@driessen1985tau] for
more details.
This game property can be checked via the 
function `isSemiConvexGame`.

### 1-convex games

A TU game $v$ with $n$ players is called **1-convex**
if for its gap function $g$ there holds
$$0 \leq g(N) \leq g(S)$$
for all coalitions $S \subseteq N$ with $S \neq \emptyset$.
Note that the 1-cover $v_{1}$ 
of a 1-convex game $v$ will always be convex.
We refer to the book by Driessen [@DriessenBook], p. 73, for
more details.
This game property can be checked via the 
function `is1ConvexGame`.

### k-convex games
k-convexity can be regarded as a generalization of 
1-convexity. 
A TU game $v$ with $n$ players is called **k-convex** if 
and only if its k-cover (see 1.2.14 of this vignette)
exists and is convex. 
We refer to section 7.1 of the book by 
Driessen [@DriessenBook] for more details on k-convex games.
This game property can be checked via the 
function `iskConvexGame` by specifying both 
a game vector $v$ and an integer $k$. We end this chapter 
with a small example inspired by the book by 
Driessen [@DriessenBook], p. 175:
```{r, collapse=TRUE}
# The following game is 2-convex
library(CoopGame)
alpha = 0.48
v=c(0,0,0,alpha,alpha,0,1)
iskConvexGame(v,2)
# The following game is not 2-convex
library(CoopGame)
alpha = 0.52
v=c(0,0,0,alpha,alpha,0,1)
iskConvexGame(v,2)
```
