---
title: "CooperativeGamesWithSpecialStructure"
author: "Jochen Staudacher"
date: "6 March 2019"
output: pdf_document
---

# Cooperative Games with Special Structure

For the convenience of the users we provide the possibility 
to generate special families of games. Our approach is 
to generate the game function as a list. 
The corresponding 
game vector can either be generated directly or 
extracted as the corresponding element of the list. Still, 
for each special family of cooperative 
game there is 

* a function for the game object (as a list)
* a function for the corresponding game vector and
* a function for the value of a coalition specified by the user.

This chapter 
gives an overview of the special families of games available 
in `CoopGame` and their usage.

## Bankruptcy games 

Bankruptcy games are studied frequently 
in game cooperative game theory, see e.g. 
the book by Gura and Maschler [@GuraMaschlerBook], the 
seminal paper by Aumann 
and Maschler [@AumannMaschler1985] from 
1985, the article by Aumann [@Aumann2002] 
from 2002 or the original paper 
by O'Neill [@ONeill1982problem] on a 
problem of rights arbitration from the Talmud from 1982.

Imagine a person dies leaving debts $d_{1}, \dots, d_{n}$ 
to $n$ creditors. However, the sum $\sum_{i=1}^{n} d_{i}$ 
of the debts is greater than the value of the estate $E$ 
of the deceased. Now we face the 
problem that the debts are mutually inconsistent as
the estate is too small in order to meet all of the debts of 
the $n$ creditors.
The game theoretic approach to bankruptcy problems 
started in 1982 with the article by O'Neill [@ONeill1982problem] 
where O'Neill defines a TU game $v$ for a 
set of creditors $N = \{ 1, \dots , n\}$, 
a debt vector $d$ of lenght $n$ and an estate $E$ as 

$$ v(C) = \max (0, E - \sum_{i \in N \backslash C} d_{i}) $$
for any coalition $C$ of creditors.

`CoopGame` provides the function `bankruptcyGame`. 
It creates a list containing all information about 
a bankruptcy game specified by the user, i.e. the 
list contains the number of players $n$, 
the vector of claims $d$, the estate $E$, 
plus the bankruptcy game vector $v$. 
Employing the function `bankruptcyGameVector` users 
can alternatively generate the corresponding game vector 
directly specifying the number $n$ of players,
the value of the estate $E$ and a vector of claims 
$d$ of length $n$. `CoopGame` will check whether 
the specification of the bankruptcy game is consistent 
in the sense that $E \leq \sum_{i \in N} d_{i}$.

We look into the usage of `bankruptcyGame` 
and `bankruptcyGameVector` by 
studying an important 
example from the Babylonian Talmud frequently 
employed in the game theory literature.

### A problem from the Babylonian Talmud
This important example from the Babylonian Talmud 
deals with a man who married three women. In their marriage
contracts the three women were promised 
the sums of 100, 200 and 300 units of money after 
the death of their husband. 
The man dies and his estate amounts to
less than 600 units. In the following we will not study 
this important example in more detail, but simply focus 
on an estate $E$ worth 300 units of money.

`CoopGame` allows us to create the corresponding 
bankruptcy game.

```{r, collapse=TRUE,eval=FALSE}
library(CoopGame)
bankruptcyGame(n=3,d=c(100,200,300),E=300)
## $n
## [1] 3
##
## $d
## [1] 100 200 300
##
## $E
## [1] 300
##
## $v
## [1]   0   0   0   0 100 200 300
```

As stated we provide the user with various ways to extract 
the corresponding game vector. One is to use 
`bankruptcyGameVector`, another to access the 
element `$v` from the `bankruptcyGame` object.

```{r, collapse=TRUE}
library(CoopGame)
# First approach
bankruptcyGameVector(n=3,d=c(100,200,300),E=300)
#
# Alternative approach
bankruptcyGame(n=3,d=c(100,200,300),E=300)$v
```

In addition, the user can also compute the value of 
any individual coalition using the function 
`bankruptcyGameValue`. For example, let us extract 
the value of the coalition of the second and third widow:

```{r, collapse=TRUE}
library(CoopGame)
bankruptcyGameValue(S=c(2,3),d=c(100,200,300),E=300)
```

We will briefly revisit bankruptcy games in 
section 5.3.2 of this vignette.
For more information on bankruptcy games and 
the contested-garment 
principle (as well as it physical 
interpretation via hydraulic rationing) we refer 
the reader to the book by
Gura and Maschler [@GuraMaschlerBook], 
the paper by Aumann and 
Maschler [@AumannMaschler1985] from 1985, 
the paper by Aumann [@Aumann2002] from 2002 
and the original paper by O'Neill [@ONeill1982problem] from 1982.

## Cost allocation games
We can look at cost allocation games in characteristic 
function form consisting of the 
set $N =\{ 1, \dots , n \}$ of agents 
(or purposes, projects or services) and the characteristic
function $c:\mathcal{P}(N) \to \mathbb{R}$ 
with $c (\emptyset) = 0$.
We are introducing the shorthand notation 
$$ c_{i} = c(\{ i \}) \quad \textrm{for} \quad i=1, \dots , n, $$
for the costs of single agents.
The connection between cost games and TU games is given by the 
associated savings 
game $v$ for $N =\{ 1, \dots , n \}$ defined by 
$$v(C) = \sum_{i \in C} c_{i} - c(C)$$
for every coalition $C$. 
Note that the associated savings game $v$ is automatically 
$0$-normalized. 
For more information on cost games 
see e.g. the article by Otten [@Otten], 
the article by Young [@YoungCostAllocation],
the overview paper by Parrachino, Zara and 
Patrone [@ParrachinoZaraPatrone], the 
paper by Tijs and Driessen [@TijsDriessen] and 
the paper by Straffin and Heaney [@StraffinHeaney].

`CoopGame` will perform computations on the corresponding 
cost-savings game. Note that if $x_{i}$, $i=1, \dots, n$, is 
the solution of the cost-savings game provided by 
`CoopGame` then the actual costs for player $i$ will 
be $y_{i} = c_{i} - x_{i}$.

`CoopGame` provides the function `costSharingGame`. 
The user needs to specify the number of players $n$ and 
a vector of costs of length $n$.
We look into the usage of `costSharingGame` 
by briefly studying an important 
example from the literature.

### The TVA problem
The TVA (Tennessee Valley Authority) problem is 
a classic example of a cost-sharing problem. We refer to 
the original article by Ransmeier 
[@ransmeier1942tennessee]
on the Tennessee Valley Authority 
from 1942 and the book by Driessen [@DriessenBook], 
pp. 1--3, for more details and 
the history of this problem. 

The following code example shows how to specify the 
TVA problem in `CoopGame`:

```{r, collapse=TRUE, eval=FALSE}
library(CoopGame)
TVACosts=c(163520,140826,250096,301607,378821,367370,412584)
(tvaCostGame <- costSharingGame(n=3, TVACosts))
## $n
## [1] 3
##
## $Costs
## [1] 163520 140826 250096 301607 378821 367370 412584
##
## $v
## [1] 0 0 0 2739 34795 23552 141858
# 
# Alternatively, generate and output only the corresponding game vector
(v <- costSharingGameVector(n=3, TVACosts))
## [1] 0 0 0 2739 34795 23552 141858
```


We will revisit the TVA problem 
in section 5.3.3 of this vignette.

## Glove games
In glove games we have a set $N = \{ 1, \dots , n\}$ 
of $n$ players and 
a disjoint union $N = L \cup R$. $L$ is the set of 
players owning one left-hand glove each and $R$ is 
the set of players owning one right-hand glove each.
The worth of a coalition $C$ equals the number of 
pairs of gloves that the members of $C$ can provide. 
In short: For every coalition $C$ there holds:
$v(C) = \min (\vert S \cap L \vert, \vert S \cap R \vert)$.
For more details on glove games we refer to 
the book by Peters [@PetersBook], p. 155--156.

`CoopGame` provides the function `gloveGame`. 
The user needs to specify the number of players $n$, the 
set $L$ of players owning one left-hand glove each and the 
set $R$ of players owning one right-hand glove each.

In the following example we compute the game vector for 
a glove game with three players and 
players 1 and 3 owning one left-hand 
glove each whereas player 2 owns the only right-hand glove. 

```{r, collapse = TRUE}
library(CoopGame)
gloveGame(n=3,L=c(1,3),R=2)$v
#
# Equivalent alternative approach
gloveGameVector(n=3,L=c(1,3),R=2)
```

## Cardinality games
For a cardinality game the value of each coalition 
is simply the number of the members of the coalition, i.e.
$v(C) = \vert C \vert$ for every coalition $C$.
In our opinion cardinality games make very good test cases. 

`CoopGame` provides the function `cardinalityGame`. 
The user only needs to specify the number of players $n$ 
in the cardinality game.

```{r, collapse=TRUE}
library(CoopGame)
cardinalityGame(4)$v
#
# Equivalent alternative approach
cardinalityGameVector(4)
```

## Weighted voting games 
We now look into decision-making and voting in 
committees and define a so-called weighted majority 
game (aka quota game, aka weighted voting game) 
for $n$ players 
as follows. Each player $j$ is assigned a weight 
$w_{j}$ (which in some situations we may also 
interpret the number of votes of a group). A law 
or motion gets passed in the committee if the 
quota $q$ is reached or exceeded, i.e.

$$ 
v(C) = \Bigg\{  
\begin{aligned}
1 & \quad \textrm{if} \quad \sum_{j \in C} w_{j} \geq q \\
0 & \quad \textrm{else.}
\end{aligned}
$$

We refer to the book by Maschler, Solan and Zamir 
[@MaschlerSolanZamir], pp. 825 -- 831, for more 
details and examples.

By definition weighted voting games 
are always simple games.
Along the lines of the book by 
Peleg and Sudhölter [@PelegSudhoelter], p. 16, 
we call a TU game $v$ **simple** if 
$v$ is monotonic and the values of all coalitions are 
$0$ or $1$, see also subsection 3.2.5 of this vignette. 


`CoopGame` provides the function `weightedVotingGame`. 
The user needs to specify the number of players $n$, 
a numeric vector of weights $w$ of length $n$ with the 
weights of the individual players and the quota $q$ 
in the weighted voting game. We provide a simple example 
with the three players.

```{r, collapse=TRUE}
library(CoopGame)
weightedVotingGame(n=3, w=c(3,5,4), q=8)$v
#
# Equivalent alternative approach
weightedVotingGameVector(n=3, w=c(3,5,4), q=8)
```

For a deeper analysis and a broader view on voting 
in committees we recommended the article by 
Peleg [@peleg2002game] from 2002.

## Weighted majority games with a single veto player
Weighted majority games with a single veto player 
are a special case of quota games.
We take our definition of a 
weighted majority game with a single veto 
player game from the book 
by Matthew O. Jackson [@JacksonBook], p. 415. 
The only winning coalitions are those containing the 
veto player $i$ and at least one other player.

$$ 
v(C) = \Bigg\{  
\begin{aligned}
1 & \quad \textrm{if} \quad \vert C \vert \geq 2 \quad 
\textrm{and} \quad i \in C\\
0 & \quad \textrm{else.}
\end{aligned}
$$


`CoopGame` provides the function `majoritySingleVetoGame`. 
The user needs to specify the number of players $n$ 
and the number of the veto player in 
the weighted majority game.

The following example produces the game vector for 
a majority game with four players and player $2$ in 
the role of the veto player.

```{r, collapse=TRUE}
library(CoopGame)
majoritySingleVetoGame(n=4, vetoPlayer=2)$v
#
# Equivalent alternative approach
majoritySingleVetoGameVector(n=4, vetoPlayer=2)
```


## Unanimity games
Unanimity games are a special family of simple games.
For a unanimity game the winning coalitions are 
exactly those coalitions 
containing the set 
$T \in \mathcal{P}(N)$, i.e. 

$$ 
v(C) = \Bigg\{  
\begin{aligned}
1 & \quad \textrm{if} \quad T \subseteq C \\
0 & \quad \textrm{else,}
\end{aligned}
$$
see Peleg and Sudhölter [@PelegSudhoelter], p. 152. 
We can understand 
$T$ as the powerful (or productive) lot among the players. 
`CoopGame` provides the function `unanimityGame`. 
The user needs to specify the number of players $n$ 
and the set of powerful players $T$ in the 
unanimity game.

The following example produces the game vector for 
a unanimity game with four players and $T$ consisting 
of players $1$ and $4$.

```{r, collapse=TRUE}
library(CoopGame)
unanimityGame(n=4, T=c(1,4))$v
#
# Equivalent alternative approach 
unanimityGameVector(n=4, T=c(1,4))
```

## Apex games
Apex games are a special family of simple games.
For an apex game there are 
two types of winning coalitions: 
The first type contains the 
so-called apex player plus at the least one other player.
The only winning coalition without the apex player 
is a coalition 
of all the other players except the apex player, see 
in particular the book by Peters [@PetersBook], p. 164, 
for more details on apex games. 
`CoopGame` provides the function `apexGame`. 
The user needs to specify the number of players $n$ 
and the number of the apex player in the apex game.
The following lines construct the game vector for an 
apex game with four players and player $2$ in the role 
of the apex player.

```{r, collapse=TRUE}
library(CoopGame)
apexGame(n=4, apexPlayer=2)$v
#
# Equivalent alternative approach 
apexGameVector(n=4, apexPlayer=2)
```

## Dictator games
Dictator games are a special case of unanimity games, 
i.e. the set $T$ in the unanimity game consists of a 
single player $i$, the dictator.
$$ 
v(C) = \Bigg\{  
\begin{aligned}
1 & \quad \textrm{if} \quad i \in  C \\
0 & \quad \textrm{else.}
\end{aligned}
$$
For more details on dictator games we refer to the 
book by Peters [@PetersBook], p. 295, and the book by 
Maschler, Solan and Zamir [@MaschlerSolanZamir], p. 764.
`CoopGame` provides the function `dictatorGame`. 
The user needs to specify the number of players $n$ 
and the number of the dictator in the dictator game.
The following lines generate the game vector for 
a dictator game with four players and player $3$ in 
the role of the dictator.

```{r, collapse=TRUE}
library(CoopGame)
dictatorGame(n=4, dictator=3)$v
#
# Equivalent alternative approach 
dictatorGameVector(n=4, dictator=3)
```


## Divide-the-dollar games
The divide-the-dollar games in `CoopGame` 
are a special family of simple games.
We take our definition of a 
divide-the-dollar game from the book 
by Matthew O. Jackson [@JacksonBook], p. 413.
For a divide-the-dollar game with $n$ players the 
winning coalitions are exactly 
those coalitions containing 
at least $n/2$ players, i.e.
$$ 
v(C) = \Bigg\{  
\begin{aligned}
1 & \quad \textrm{if} \quad \vert C \vert \geq n/2 \\
0 & \quad \textrm{else.}
\end{aligned}
$$

`CoopGame` provides the function `divideTheDollarGame`. 
The user only needs to specify the number of players $n$ 
in the divide-the-dollar game.
We end this chapter with the game vector for 
a divide-the-dollar game with three players.

```{r, collapse=TRUE}
library(CoopGame)
divideTheDollarGame(3)$v
#
# Equivalent alternative approach
divideTheDollarGameVector(3)
```
